\ \\ \LARGE\bf Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*

### 2014-01-17 tex/cec14/article.tex
[tex]

In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
First, \lets 
describe a programming language, 
in which the GP algorithm generates individual programs --- the so called \lterms.

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

@@

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).

\begin{definition}
Let $A$ be set of {\it atomic type names}. 
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}
\end{definition}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$. 
$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
combining the previously defined \textit{\lterms} and \textit{types} together.

\begin{definition}\begin{enumerate}
 \item 	Let $\Lambda$ be set of {\it \lterms}. 
	Let $\mathbb{T}$ be set of {\it types}.       
	A {\it statement} $M : \sigma$ is a pair 
	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
	Statement $M : \sigma$ is vocalized as 
	{\it "$M$ has type $\sigma$"}.
	The term $M$ is called the {\it subject} of the 
	statement $M : \sigma$.
 \item A \textit{declaration} is a statement 
 $x : \sigma$ where $x \in V \cup C$.

 \item A \textit{context} 
 is set of declarations with distinct variables as subjects.
\end{enumerate}
\end{definition}

Context is a basic type theoretic concept suitable as a typed alternative
for terminal and function set in standard GP. 
Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
such that $\Gamma$ does not contain any declaration with $x$ as subject.
We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation 
\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
if it can be produced by the following rules.
\begin{align*}
x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma,x:\sigma}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}
\end{definition}

Our goal in term generation is to produce terms $M$
for a given pair $\ul{\tau}{\Gamma}$
such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	
Let $\T$ be set of types.
Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
Then $\Lpr$ is set of 
\textit{unfinished  \lterms} defined inductively as follows.	
\begin{align*}
\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
\textit{~~~~~~~~(Unfinished leaf)}\\
x   \in V \cup C  &\then x     \in \Lpr \\
M,N \in \Lpr   &\then (M~N) \in \Lpr 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

@@

\Lets now discuss three such search strategies.

\subsubsection{Systematic strategy}

If we use trivial strategy that returns all the inputs, 
then the algorithm systematically generates 
first $n$ smallest lambda terms in their
\textit{long normal form}.

\subsubsection{Ramped half-and-half strategy}
is generalization of the standard ramped half-and-half method described 
in \cite{koza92}. If applied to context satisfying closure requirement
(all constants/variables are of the same type, all functions are operations
on this type), it will behave in the same way as the standard method.
The internal state of this strategy consists of two variables.
It is the only one strategy described here that uses an internal state.

@@

This strategy returns precisely one randomly
(uniformly) selected  element from 
the \textit{selection subset} of input set
(or zero elements if the input set is empty). 
The \textit{selection subset} 
to select from is determined by $depth, d$ and $isFull$.
The $depth$ parameter is the depth (in the term tree) 
of the unfinished leaf that was expanded.
Those elements of input set whose newly added subtree contains one ore more 
unfinished leafs are regarded as \textit{non-terminals}, whereas 
those whose newly added subtree contains no unfinished leaf are regarded as 
\textit{terminals}.
If $depth = 0$, then the subset to select from is  
set of all \textit{non-terminals} of the input set.
If $depth = d$, then the subset to select from is
set of all \textit{terminals} of the input set.
In other cases of $depth$ it depends on value of $isFull$.
If $isFull = true$, then the subset to select from is 
set of all \textit{non-terminals} of the input set.
If $isFull = false$, then the subset to select from is 
the whole input set.

\subsubsection{Geometric strategy}

We can see those two previous strategies as two extremes on the spectrum of 
possible strategies. 
\textit{Systematic strategy} filters no successor state thus performing
exhaustive search resulting in discovery of $n$ smallest terms in one run.
On the other hand, \textit{ramped half-and-half strategy} filters 
all but one successor states resulting in degradation of 
the priority queue into "fancy variable".
\textit{Geometric strategy} is simple yet fairly effective term generating 
strategy somewhere in the middle of this spectrum.
It is parameterized by parameter $q \in (0,1)$, its default well-performing 
value is $q = 0.75$.
For each element of the input set 
it is probabilistically decided whether
it will be returned or omitted. A probability $p$ of returning is
same for all elements, but depends on the $depth$, 
which is defined in the same way as in previous strategy. 
It is computed as follows.
$$ p = q^{depth} $$
This formula is motivated by idea that it is important to
explore all possible root symbols, but as the $depth$ 
increases it becomes less "dangerous" to omit 
an exploration branch. 
We can see this by considering that this strategy results in
somehow forgetful A* search.
With each omission we make the search space smaller. But with
increasing depth these omissions have smaller impact on the search space,
i.e., they cut out lesser portion of the search space.
Another slightly esoteric argument supporting this formula is that "root 
parts" of a program usually stand for more crucial parts
with radical impact on global behavior of a program, 
whereas "leaf parts" of a program usually
stand for less important local parts (e.g. constants).  
This strategy also plays nicely with the idea that 
"too big trees should be killed".

@@

\subsection{Simple symbolic regression}
\textit{Simple Symbolic Regression} is a problem described
in \cite{koza92}. Objective of this problem is to 
find a function $f(x)$ that fits a sample
of twenty given points. The target function is 
function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
Desired type of generated programs $\tau$ and 
building blocks context $\Gamma$ are following.
\begin{align*}
\tau = \Real \ar &\Real\\
\Gamma = \{
  (+)  &: \Real \ar \Real \ar \Real    ,
  (-)   : \Real \ar \Real \ar \Real    ,
  (*)   : \Real \ar \Real \ar \Real    ,
  rdiv  : \Real \ar \Real \ar \Real    ,\\
  sin  &: \Real \ar \Real              ,
  cos   : \Real \ar \Real              ,
  exp   : \Real \ar \Real              , 
  rlog  : \Real \ar \Real              \}
\end{align*}
where
\noindent
\begin{minipage}{.5\linewidth}
\begin{align*}
rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
p/q & \mbox{otherwise } \end{cases}  
\end{align*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
\begin{align*}
rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
log(\vert x\vert) & \mbox{otherwise}. \end{cases}
\end{align*}
\end{minipage}

\begin{figure}[h!]
  \caption{Performance curves for Simple symbolic regression.}
  \centering
    \includegraphics[scale=0.135]{imgs/SSR_FINAL.png}
\end{figure}
Fitness function is computed as follows.
$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
considered as a correct individual.

@@

\begin{align*}
\tau = An&tAction\\
\Gamma = \{~~
  l    &: AntAction                              ,
  r    : AntAction                               ,
  m    : AntAction                               ,\\
  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
\end{align*}

@@

\begin{figure}[h!]
  \caption{Performance curves for Artificial ant problem.}
  \centering
    \includegraphics[scale=0.135]{imgs/ANT_FINAL.png}
\end{figure}

@@

Previous two problems were instances of standard GP, i.e. their building symbols
obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
The even-parity function is a boolean function taking as inputs $N$
boolean values and returning \textit{True} if an even number of inputs 
are \textit{True}. For odd number it returns \textit{False}.
This problem has been used by many researchers
as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
where is presented approach to evolve recursive and modular programs
by use of higher-order functions and $\lambda$-abstractions.
We use very similar set of \textit{building symbols} as in \cite{yu01}. 
\begin{align*}
\tau = [Boo&l] \ar Bool\\
\Gamma = \{
  and   &: Bool \ar Bool \ar Bool                              ,
  or    : Bool \ar Bool \ar Bool                              ,\\
  nand  &: Bool \ar Bool \ar Bool                              ,
  nor    : Bool \ar Bool \ar Bool                              ,\\
  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
  head' &: [Bool] \ar Bool                                   ,
  tail' : [Bool] \ar [Bool]                              \}
\end{align*}
\begin{figure}[h!]
  \caption{Performance curves for Even parity problem.}
  \centering
    \includegraphics[scale=0.135]{imgs/EP_FINAL.png}
\end{figure}

@@

\begin{thebibliography}{22}

\bibitem{koza92}
  TODO. Todo,
  \emph{Dát sem místo toho ten bibtex}.
  MIT Press, Cambridge, MA,
  2014.

\bibitem{koza92}
  John R. Koza,
  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
  MIT Press, Cambridge, MA,
  1992.

\bibitem{koza05}
  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
  Springer, 2005. ISBN 978-0-387-26417-2

\bibitem{fg}
 Riccardo Poli, William B. Langdon, Nicholas F. McPhee
 \emph{A Field Guide to Genetic Programming}.
 Lulu Enterprises, UK Ltd, 2008.

\bibitem{yu01}
  T. Yu. 
  \emph{Hierachical processing for evolving recursive and modular 
        programs using higher order functions and lambda abstractions}. 
  Genetic Programming and Evolvable Machines,
  2(4):345–380, December 2001. ISSN 1389-2576.

\bibitem{montana95}
D. J. Montana. 
\emph{Strongly typed genetic programming.} 
Evolutionary Computation, 3(2): 199–230, 1995.
%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha

\bibitem{haynes96}
T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
\emph{Type inheritance in strongly typed genetic programming.} 
In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
in Genetic Programming 2, chapter 18, pages 359–376.
MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.

\bibitem{olsson94}
J. R. Olsson. 
\emph{Inductive functional programming using incremental program 
transformation and Execution of logic programs by 
iterative-deepening A* SLD-tree search.} 
Dr scient thesis, University of Oslo, Norway, 1994.

\bibitem{kes}
Forrest Briggs, Melissa O’Neill.
\emph{Functional Genetic Programming and Exhaustive
Program Search with Combinator Expressions.}
International Journal of Knowledge-based and Intelligent Engineering Systems,
Volume 12 Issue 1, Pages 47-68, January 2008.

\bibitem{barendregt84}
H. P. Barendregt,
\emph{The Lambda Calculus: its Syntax and Semantics}, 
revised ed., North-Holland, 1984.

\bibitem{barendregt92}
H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
\emph{Lambda Calculi with Types.} 
Handbook of Logic in Computer Science, 1992.

\bibitem{barendregt10}

  Henk Barendregt, Wil Dekkers, Richard Statman,
  \emph{Lambda Calculus With Types}.
  Cambridge University Press,
  2010. 
  %URL \url{http://www.cs.ru.nl/~henk/book.pdf}.

\bibitem{jones87}
Simon Peyton Jones. 
\emph{The Implementation of Functional Programming Languages}. 
Prentice Hall, 1987.

\bibitem{AIAMA}
	Stuart J. Russell, Peter Norvig,
	\emph{Artificial Intelligence: A Modern Approach}.
	Pearson Education,
	2003.

\end{thebibliography}

>>>>

In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
First, \lets describe a programming language, 
in which the GP algorithm generates individual programs --- the so called \lterms.\\

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}
~\\

@@

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).\\

\begin{definition}
Let $A$ be set of {\it atomic type names}. 
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}
\end{definition}
~\\

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$. 
$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
combining the previously defined \textit{\lterms} and \textit{types} together. \\

\begin{definition}~

\begin{enumerate}
 \item 	Let $\Lambda$ be set of {\it \lterms}. 
	Let $\mathbb{T}$ be set of {\it types}.       
	A {\it statement} $M : \sigma$ is a pair 
	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
	Statement $M : \sigma$ is vocalized as 
	{\it "$M$ has type $\sigma$"}.
	The term $M$ is called the {\it subject} of the 
	statement $M : \sigma$.
 \item A \textit{declaration} is a statement 
 $x : \sigma$ where $x \in V \cup C$.

 \item A \textit{context} 
 is set of declarations with distinct variables as subjects.
\end{enumerate}
\end{definition}
%~\\

Context is a basic type theoretic concept suitable as a typed alternative
for terminal and function set in standard GP. 
Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
such that $\Gamma$ does not contain any declaration with $x$ as subject.
We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.\\

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation 
\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
if it can be produced by the following rules.
\begin{align*}
x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma,x:\sigma}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}
\end{definition}
~\\

Our goal in term generation is to produce terms $M$
for a given pair $\ul{\tau}{\Gamma}$
such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.\\

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	
Let $\T$ be set of types.
Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
Then $\Lpr$ is set of 
\textit{unfinished  \lterms} defined inductively as follows.	
\begin{align*}
\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
\textit{~~~~~~~~(Unfinished leaf)}\\
x   \in V \cup C  &\then x     \in \Lpr \\
M,N \in \Lpr   &\then (M~N) \in \Lpr 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}
~\\

@@

\Lets now discuss three such search strategies.\\

%\subsubsection{Systematic strategy}
\textit{Systematic strategy}:~~ 
If we use trivial strategy that returns all the inputs, 
then the algorithm systematically generates 
first $n$ smallest lambda terms in their
\textit{long normal form}.\\

%\subsubsection{Ramped half-and-half strategy}
\textit{Ramped half-and-half strategy}:~~
This is generalization of the standard ramped half-and-half method described 
in \cite{koza92}. If applied to context satisfying closure requirement
(all constants/variables are of the same type, all functions are operations
on this type), it will behave in the same way as the standard method.
The internal state of this strategy consists of two variables.
It is the only one strategy described here that uses an internal state.

@@

This strategy returns precisely one randomly
(uniformly) selected  element from 
the \textit{selection subset} of input set
(or zero elements if the input set is empty). 
The \textit{selection subset} 
to select from is determined by $depth, d$ and $isFull$.
The $depth$ parameter is the depth (in the term tree) 
of the unfinished leaf that was expanded.
Those elements of input set whose newly added subtree contains one ore more 
unfinished leafs are regarded as \textit{non-terminals}, whereas 
those whose newly added subtree contains no unfinished leaf are regarded as 
\textit{terminals}.
If $depth = 0$, then the subset to select from is  
set of all \textit{non-terminals} of the input set.
If $depth = d$, then the subset to select from is
set of all \textit{terminals} of the input set.
In other cases of $depth$ it depends on value of $isFull$.
If $isFull = true$, then the subset to select from is 
set of all \textit{non-terminals} of the input set.
If $isFull = false$, then the subset to select from is 
the whole input set.\\

%\subsubsection{Geometric strategy}
\textit{Geometric strategy}:~~
We can see those two previous strategies as two extremes on the spectrum of 
possible strategies. 
\textit{Systematic strategy} filters no successor state thus performing
exhaustive search resulting in discovery of $n$ smallest terms in one run.
On the other hand, \textit{ramped half-and-half strategy} filters 
all but one successor states resulting in degradation of 
the priority queue into "fancy variable".
\textit{Geometric strategy} is simple yet fairly effective term generating 
strategy somewhere in the middle of this spectrum.
It is parameterized by parameter $q \in (0,1)$, its default well-performing 
value is $q = 0.75$.
For each element of the input set 
it is probabilistically decided whether
it will be returned or omitted. A probability $p$ of returning is
same for all elements, but depends on the $depth$, 
which is defined in the same way as in previous strategy. 
It is computed as follows.
$$ p = q^{depth} $$
This formula is motivated by idea that it is important to
explore all possible root symbols, but as the $depth$ 
increases it becomes less "dangerous" to omit 
an exploration branch. 
We can see this by considering that this strategy results in
somehow forgetful A* search.
With each omission we make the search space smaller. But with
increasing depth these omissions have smaller impact on the search space,
i.e., they cut out lesser portion of the search space.
Another slightly esoteric argument supporting this formula is that "root 
parts" of a program usually stand for more crucial parts
with radical impact on global behavior of a program, 
whereas "leaf parts" of a program usually
stand for less important local parts (e.g. constants).  
This strategy also plays nicely with the idea that 
"too big trees should be killed".

@@

\subsection{Simple symbolic regression}
\textit{Simple Symbolic Regression} is a problem described
in \cite{koza92}. Objective of this problem is to 
find a function $f(x)$ that fits a sample
of twenty given points. The target function is 
function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
Desired type of generated programs $\tau$ and 
building blocks context $\Gamma$ are following.
\begin{align*}
\tau = \Real \ar &\Real\\
\Gamma = \{
  (+)  &: \Real \ar \Real \ar \Real    ,
  (-)   : \Real \ar \Real \ar \Real    ,
  (*)   : \Real \ar \Real \ar \Real    ,\\
  rdiv &: \Real \ar \Real \ar \Real    ,
  sin   : \Real \ar \Real              ,
  cos   : \Real \ar \Real              ,\\
  exp  &: \Real \ar \Real              , 
  rlog  : \Real \ar \Real              \}
\end{align*}
where

\noindent
\begin{minipage}{.5\linewidth}
\begin{align*}
rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
p/q & \mbox{otherwise } \end{cases}  
\end{align*}
\end{minipage}%

\begin{minipage}{.5\linewidth}
\begin{align*}
rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
log(\vert x\vert) & \mbox{otherwise}. \end{cases}
\end{align*}
\end{minipage}

\begin{figure}[h!]
  \caption{Performance curves for Simple symbolic regression.}
  \centering
    \includegraphics[scale=0.1]{imgs/SSR_FINAL.png}
\end{figure}
Fitness function is computed as follows.
$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
considered as a correct individual.

@@

\begin{align*}
\tau = An&tAct\\
\Gamma = \{~~
  l    &: AntAct                              ,
  r    : AntAct                               ,
  m    : AntAct                               ,\\
  ifa  &: AntAct \ar AntAct \ar AntAct  ,\\
  p2   &: AntAct \ar AntAct \ar AntAct  ,\\
  p3   &: AntAct \ar AntAct \ar AntAct \ar AntAct  \}
\end{align*}

@@

\begin{figure}[h!]
  \caption{Performance curves for Artificial ant problem.}
  \centering
    \includegraphics[scale=0.1]{imgs/ANT_FINAL.png}
\end{figure}

@@

Previous two problems were instances of standard GP, i.e. their building symbols
obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
The even-parity function is a boolean function taking as inputs $N$
boolean values and returning \textit{True} if an even number of inputs 
are \textit{True}. For odd number it returns \textit{False}.
This problem has been used by many researchers
as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
where is presented approach to evolve recursive and modular programs
by use of higher-order functions and $\lambda$-abstractions.
We use very similar set of \textit{building symbols} as in \cite{yu01}. 
\begin{align*}
\tau = [Boo&l] \ar Bool\\
\Gamma = \{
  and   &: Bool \ar Bool \ar Bool                              ,
  or    : Bool \ar Bool \ar Bool                              ,\\
  nand  &: Bool \ar Bool \ar Bool                              ,
  nor    : Bool \ar Bool \ar Bool                              ,\\
  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
  head' &: [Bool] \ar Bool                                   ,
  tail' : [Bool] \ar [Bool]                              \}
\end{align*}
\begin{figure}[h!]
  \caption{Performance curves for Even parity problem.}
  \centering
    \includegraphics[scale=0.1]{imgs/EP_FINAL.png}
\end{figure}

@@

%\nocite{*}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,evogp}
